Resonant Scattering in a Uniform Sphere with Large Optical Depth

### 2021-07-07 ms.tex
Fixing sign on eq 54
adding discussion of the scaling of Pnm with m

Let us define the damping time to be $\gamma \equiv i\omega$, which is real-valued and positive for damped solutions. There are certain values of $\gamma$ for which a resonant response is produced which we will call $\gamma_{nm}$. Here, $m$ is an index that represents individual frequency eigenmodes, as opposed to $n$ which represents the spatial mode number. Near these frequencies, the solution will have the form 
\be
J(n,\sigma,-i\gamma) & \simeq \frac{ J_{nm}(\sigma) }{\gamma_{nm} - \gamma} + {\rm terms\ which\ vary\ slowly\ in\ \gamma}.
\ee
Obtaining the values of these eigenfrequencies and their corresponding eigenfunctions $J_{nm}(\sigma)$ allows us to obtain the contribution of this mode to the total solution via contour integration. 
\be
J(r,\sigma,t) & = &  -j_0(\kappa_n r)  \int \frac{d\gamma}{2\pi i} e^{-\gamma t}J(n,\sigma,-i\gamma)
\nonumber \\ & \simeq & j_0(\kappa_n r)  \int \frac{d\gamma}{2\pi i} e^{-\gamma t} \frac{ J_{nm}(\sigma) }{\gamma - \gamma_{nm}} 
\nonumber \\ & \simeq & j_0(\kappa_n r)  J_{nm}(\sigma) e^{-\gamma_{nm}t}.
\ee
Here we have closed the integral along the imaginary $\gamma$ axis by a semi-circle at infinity, using the residue theorem to obtain the answer. The contribution from the semi-circle at infinity is zero. Summing over all spatial eigenmodes $n$ and over all temporal eigenmodes $m$ for a given $n$, we get the final form of the result
\be \label{eq:Jrsigmat}
J(r,\sigma,t) & = & \sum_{n=1}^\infty j_0(\kappa_n r)  \sum_m J_{nm}(\sigma) e^{-\gamma_{nm}t}.
\ee
Beginning with Eq. \ref{eq:Jrsigmat}, we use the derivative
\be
\frac{dj_0(\kappa_n r)}{dr} \rfloor_R & =& \frac{d}{dr}  \rfloor_R \left( \frac{\sin(\kappa_n r)}{\kappa_n r} \right)
=  \left( \frac{\cos(\kappa_n R)}{R} - \frac{\sin(\kappa_n R)}{\kappa_n R^2} \right)\rfloor_R
\nonumber \\ & = & \frac{(-1)^n}{R},
\ee
to obtain the flux at the surface, which is
\be
F(R,\sigma,t) & =& - \frac{4\pi}{3k\phi} \frac{dJ(R,\sigma,t)}{dr} 
= - \frac{4\pi}{3k\phi R}  \sum_{nm} (-1)^n J_{nm}(\sigma) e^{-\gamma_{nm}t}.
\ee
Multiplying by $4\pi R^2$ gives the energy per time per frequency
\be
\frac{dE}{dtd\nu} & = & - \frac{16\pi^2 R}{3k\phi}  \sum_{nm} (-1)^n J_{nm}(\sigma) e^{-\gamma_{nm}t}.
\label{eq:dEdtdnu}
\ee
Integrating this over time yields a factor $-1/\gamma_{nm}$, and over $d\nu = \sqrt{3/2} \Delta^2 \phi d\sigma$, a ``sum rule" results
\be
1 & = &  \sqrt{ \frac{3}{2} } \frac{16\pi^2R\Delta^2}{3kE} \sum_{nm} (-1)^n \gamma_{nm}^{-1} \int d\sigma J_{nm}(\sigma).
\ee
This non-trivial expression provides a check on the values of $\gamma_{nm}$ and the functions $J_{nm}(\sigma)$. This expression can also be written as
\be
1 & =& \sum_{nm} P_{nm},
\label{eq:sumrule}
\ee
where the contribution of each eigenfunction is
\be
P_{nm} & \equiv & \sqrt{ \frac{3}{2} } \frac{16\pi^2R\Delta^2}{3kE}  (-1)^n \gamma_{nm}^{-1} \int d\sigma J_{nm}(\sigma).
\ee
These coefficients can have either sign depending on the value of $n$.

@@

Integrating $dE/dt d\nu$ from Eq. \ref{eq:dEdtdnu} over all times should give a distribution for the emitted frequencies. It is expected that this time-integrated spectrum of the response to an impulse will agree well with the solution for the $J=0$ steady-state spectrum (Eq. \ref{eq:H0surf}). Dividing by the energy $E$, we find the probability distribution
\be \label{eq:spectrum}
P(\nu) & = &  -\frac{16\pi^2 R}{3k\phi E}  \sum_{nm} (-1)^n \gamma_{nm}^{-1} J_{nm}(\sigma).
\ee
Integrating over $\nu$ gives unity if the sum rule in Eq. \ref{eq:sumrule} is true. In Fig. \ref{fig:steadystate}, the resulting spectrum is shown for a sum up to $n=20$ and $m=500$ (labelled ``Time-integrated'') as compared with two analytic solutions: the steady-state $J=0$ solution with no time dependence (``Steady State'') and the expected result for summing \textit{only} spatial modes in the eigenfunction expansion with the time dependence ignored (``Partial Sum''). This graphic illustrates the physics modeled by each variable in the expansion. For example, the spatial modes contribute to the solutions' accuracy in the core of the line. If more spatial modes were calculated, the agreement between the lines extends further into the core. If additional frequency modes are calculated, faster-oscillating terms are incorporated into the sum over the imaginary frequency variable $\gamma$ which create more perfect cancellations with the lower-order terms, reducing the ``ringing'' seen in the time-integrated spectrum. Modeling the physics deep into the line core could have an impact on the accuracy of the escape time distribution, but this would only affect late times since core photons are trapped for a longer duration before escaping.

>>>>

Let us define the damping time to be $\gamma \equiv i\omega$, which is real-valued and positive for damped solutions. There are certain values of $\gamma$ for which a resonant response is produced which we will call $\gamma_{nm}$. Here, $m$ is an index that represents individual frequency eigenmodes, as opposed to $n$ which represents the spatial mode number. Near these frequencies, the solution will have the form 
\be
J(n,\sigma,-i\gamma) & \simeq \frac{ J_{nm}(\sigma) }{\gamma_{nm} - \gamma} + {\rm terms\ which\ vary\ slowly\ in\ \gamma}.
\ee
Obtaining the values of these eigenfrequencies and their corresponding eigenfunctions $J_{nm}(\sigma)$ allows us to obtain the contribution of this mode to the total solution via contour integration. 
\be
J(r,\sigma,t) & = &  -j_0(\kappa_n r)  \int \frac{d\gamma}{2\pi i} e^{-\gamma t}J(n,\sigma,-i\gamma)
\nonumber \\ & \simeq & j_0(\kappa_n r)  \int \frac{d\gamma}{2\pi i} e^{-\gamma t} \frac{ J_{nm}(\sigma) }{\gamma - \gamma_{nm}} 
\nonumber \\ & \simeq & j_0(\kappa_n r)  J_{nm}(\sigma) e^{-\gamma_{nm}t}.
\ee
Here we have closed the integral along the imaginary $\gamma$ axis by a semi-circle at infinity, using the residue theorem to obtain the answer. The contribution from the semi-circle at infinity is zero. Summing over all spatial eigenmodes $n$ and over all temporal eigenmodes $m$ for a given $n$, we get the final form of the result
\be \label{eq:Jrsigmat}
J(r,\sigma,t) & = & \sum_{n=1}^\infty j_0(\kappa_n r)  \sum_m J_{nm}(\sigma) e^{-\gamma_{nm}t}.
\ee
Beginning with Eq. \ref{eq:Jrsigmat}, we use the derivative
\be
\frac{dj_0(\kappa_n r)}{dr} \rfloor_R & =& \frac{d}{dr}  \rfloor_R \left( \frac{\sin(\kappa_n r)}{\kappa_n r} \right)
=  \left( \frac{\cos(\kappa_n R)}{R} - \frac{\sin(\kappa_n R)}{\kappa_n R^2} \right)\rfloor_R
\nonumber \\ & = & \frac{(-1)^n}{R},
\ee
to obtain the flux at the surface, which is
\be
F(R,\sigma,t) & =& - \frac{4\pi}{3k\phi} \frac{dJ(R,\sigma,t)}{dr} 
= - \frac{4\pi}{3k\phi R}  \sum_{nm} (-1)^n J_{nm}(\sigma) e^{-\gamma_{nm}t}.
\ee
Multiplying by $4\pi R^2$ gives the energy per time per frequency
\be
\frac{dE}{dtd\nu} & = & - \frac{16\pi^2 R}{3k\phi}  \sum_{nm} (-1)^n J_{nm}(\sigma) e^{-\gamma_{nm}t}.
\label{eq:dEdtdnu}
\ee
Integrating this over time yields a factor $-1/\gamma_{nm}$, and over $d\nu = \sqrt{3/2} \Delta^2 \phi d\sigma$, a ``sum rule" results
\be
1 & = &  \sqrt{ \frac{3}{2} } \frac{16\pi^2R\Delta^2}{3kE} \sum_{nm} (-1)^n \gamma_{nm}^{-1} \int d\sigma J_{nm}(\sigma).
\ee
This non-trivial expression provides a check on the values of $\gamma_{nm}$ and the functions $J_{nm}(\sigma)$. This expression can also be written as
\be
1 & =& \sum_{nm} P_{nm},
\label{eq:sumrule}
\ee
where the contribution of each eigenfunction is
\be
P_{nm} & \equiv & \sqrt{ \frac{3}{2} } \frac{16\pi^2R\Delta^2}{3kE}  (-1)^n \gamma_{nm}^{-1} \int d\sigma J_{nm}(\sigma).
\ee
These coefficients can have either sign depending on the value of $n$. The size of each contribution scales roughly as $0.5/(m+1/8)^{2/3}$, with a very weak dependence on $n$. This indicates the need for a large number of $n$ and $m$ to converge, i.e., it takes roughly ten times as many $m$ modes to reduce the size of $P_{nm}$ by a factor of ${\sim}$5. Numerical convergence would be obtained only when the size of additional $P_{nm}$ are negligible in the sum.

@@

Integrating $dE/dt d\nu$ from Eq. \ref{eq:dEdtdnu} over all times should give a distribution for the emitted frequencies. It is expected that this time-integrated spectrum of the response to an impulse will agree well with the solution for the $J=0$ steady-state spectrum (Eq. \ref{eq:H0surf}). Dividing by the energy $E$, we find the probability distribution
\be \label{eq:spectrum}
P(\nu) & = &  \frac{16\pi^2 R}{3k\phi E}  \sum_{nm} (-1)^n \gamma_{nm}^{-1} J_{nm}(\sigma).
\ee
Integrating over $\nu$ gives unity if the sum rule in Eq. \ref{eq:sumrule} is true. In Fig. \ref{fig:steadystate}, the resulting spectrum is shown for a sum up to $n=20$ and $m=500$ (labelled ``Time-integrated'') as compared with two analytic solutions: the steady-state $J=0$ solution with no time dependence (``Steady State'') and the expected result for summing \textit{only} spatial modes in the eigenfunction expansion with the time dependence ignored (``Partial Sum''). This graphic illustrates the physics modeled by each variable in the expansion. For example, the spatial modes contribute to the solutions' accuracy in the core of the line. If more spatial modes were calculated, the agreement between the lines extends further into the core. If additional frequency modes are calculated, faster-oscillating terms are incorporated into the sum over the imaginary frequency variable $\gamma$ which create more perfect cancellations with the lower-order terms, reducing the ``ringing'' seen in the time-integrated spectrum. Modeling the physics deep into the line core could have an impact on the accuracy of the escape time distribution, but this would only affect late times since core photons are trapped for a longer duration before escaping. This was the motivation for choosing a comparatively low number of spatial eigenmodes with respect to the number of frequency eigenmodes calculated.
